---
title: "Subsetting and Flow Control Statements (under construction)"
comment: "*change the flow of your code*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    theme: flatly
    df_print: paged

always_allow_html: true
---

```{r, echo = FALSE}
# don't name this setup!
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      message = FALSE,
                      warning = FALSE)
```

# Objectives

This section contains a basic overview of R's subsetting syntax (`[[`, `[`, and `$`) and flow control functions (`if`, `ifelse`, and `switch`), and how they can be used to control data workflows. 

## Subset vectors {.tabset}

Vectors are the fundamental data type in R. Below we assign five numbers to the vector `vec`. 

```{r num_vec}
num_vec <- c(2, 9, 4, 3, 7)
num_vec
```

These five numbers exist in five different positions within the vector, and we can subset them using a bracket with the numerical 

### subset using `[`

Here are the numbers located at positions `[1]` and position `[5]`

```{r subset-by-position}
num_vec[1]
num_vec[5]
```

### subset using `$`

## Subset data frames/tibbes {.tabset}

## Subset lists {.tabset}

## `if` {.tabset}

The `if` statement can be used to 

```{r if-statement}

```
