---
title: "Creating Data in R"
comment: "*introduction to data generation and simulation*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: show
    theme: flatly
    df_print: paged

always_allow_html: true
---

```{r , echo = FALSE, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      error = TRUE)
library(tidyverse)
library(skimr)
```


# Objectives

This lesson outlines some examples for creating data in R. Being able to generate your own datasets in R makes it easier to test code and create reproducible examples.  

# Materials

The slides for this presentation are [here]().

The RStudio.Cloud project [here]().


Read more about data simulation [here](https://rviews.rstudio.com/2019/07/01/imagine-your-data-before-you-collect-it/) and [here](https://rviews.rstudio.com/2020/09/09/fake-data-with-r/) on the RStudio blog.

# Load packages

Great packages for generating and simulating data are `fakeR`, `fabricatr`, `charlatan` and `fakir`.

```{r packages}
# install.packages(c("fakeR", "fabricatr", 
#                    "charlatan", "fakir", 
#                    "tidyverse"))
library(fakeR)
library(fabricatr)
library(charlatan)
library(fakir)
library(tidyverse)
```



# Import data

We'll cover two methods for importing data into RStudio.
