---
title: "Advanced Data Visualization"
comment: "*extending ggplot2*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: show
    theme: flatly
    df_print: paged

always_allow_html: true
---


```{r , include=FALSE}
library(tidyverse)
library(skimr)
library(lubridate)
library(here)
library(janitor)
library(socviz)
library(ggrepel)
library(covdata)
library(showtext)
library(hrbrthemes)
# create data folder
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.path = "img/",
                      tidy = FALSE)
# set width
options(width = 60, max.print = 60)
```

# Objectives

Give your graphs a custom look using `ggplot` extension packages.

# Materials

- Slides:

- RStudio Project:

# Previous lessons

All of the exercises and lessons are available [here](https://mjfrigaard.github.io/r-meetup-tutorials/), but you can also read more about [`ggplot2`](https://ggplot2.tidyverse.org/index.html) on the tidyverse website, and in the [Data Visualisation](https://r4ds.had.co.nz/data-visualisation.html) chapter of R for Data Science.

# Load the `tidyverse`

The main packages we're going to use are `dplyr`, `tidyr`, and `ggplot2`. These are all part of the `tidyverse`, so we'll import this package below:

```{r packages, eval=FALSE}
install.packages("tidyverse")
library(tidyverse)
```

# `ggplot2` extension packages

```{r ggpl2-packages, message=FALSE, warning=FALSE}
library(ggforce)
library(ggbeeswarm)
library(ggfittext)
library(ggthemes)
```

