---
title: "Advanced Data Visualization"
comment: "*extending ggplot2*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: show
    theme: flatly
    df_print: paged

always_allow_html: true
---


```{r , include=FALSE}
library(tidyverse)
library(skimr)
library(lubridate)
library(here)
library(janitor)
library(socviz)
library(ggrepel)
library(covdata)
library(showtext)
library(hrbrthemes)
library(tidytuesdayR)
library(Lahman)
library(starwarsdb)
library(dplyr)
# create data folder
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.path = "img/",
                      tidy = FALSE)
# set width
options(width = 60, max.print = 60)
```

# Objectives

Give your graphs a custom look using `ggplot` extension packages.

# Materials

- Slides: *this lesson does not slides.*

- RStudio Project: *this lesson does not have an RStudio.Cloud project.*

# Previous lessons

All of the exercises and lessons are available [here](https://mjfrigaard.github.io/r-meetup-tutorials/), but you can also read more about [`ggplot2`](https://ggplot2.tidyverse.org/index.html) on the tidyverse website, and in the [Data Visualisation](https://r4ds.had.co.nz/data-visualisation.html) chapter of R for Data Science.

# Load the `tidyverse`

The main packages we're going to use are `dplyr`, `tidyr`, and `ggplot2`. These are all part of the `tidyverse`, so we'll import this package below:

```{r packages, eval=FALSE}
install.packages("tidyverse")
library(tidyverse)
```

# `ggplot2` extension packages

`ggplot2` has an extensive list of user-written packages for just about any data visualization you can think of, and we're only going to cover a few in this lesson. Check out more [here](https://exts.ggplot2.tidyverse.org/gallery/). 

```{r ggpl2-packages, message=FALSE, warning=FALSE}
library(ggtext)
library(ggfittext)
library(ggdist)
library(ggbeeswarm)
library(plotly)
library(ggthemes)
library(wesanderson)
library(gganimate)
```

# Data 

We're going to be using data from the following packages: 

```{r data-packages, eval=FALSE}
library(tidytuesdayR)
library(Lahman)
library(starwarsdb)
library(dplyr)
```


## Advanced Text Annotations

In the [previous lesson](https://mjfrigaard.github.io/r-meetup-tutorials/gg01_data-viz-as-comm.html#8_Adding_Text_to_Graphs), we covered how to annotate your graphs with `ggplot2::annotate()`, and how to label data points with `ggrepel`.

In this section, we're going to cover additional text options with `ggtext` and `ggfittext`.

```{r text-packages}
library(ggtext)
library(ggfittext)
```

We're going to be visualizing the relationship between height and weight for the various species in the `People` dataset from the `Lahman` package.  



```{r BBPData}
BBPData <- Lahman::People %>% 
  dplyr::select(nameGiven, height, weight)
BBPData
```


## Advanced Distributions

`ggdist`

## Advanced Colors

`wesanderson`

## Advanced Interactivity 

`plotly`

## Advanced Animations

`gganimate`

