---
title: "Displaying Tables With R Markdown (under construction)"
comment: "*table packages for reproducible reports*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    theme: flatly
    df_print: paged

always_allow_html: true
---

```{r , include=FALSE}
library(tidyverse)
library(ipcwswitch)
library(skimr)
# create data folder
# folders
folders <- c("code", "data", "img", "docs", "pdfs")
purrr::map(.x = folders, .f = fs::dir_create)
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      tidy = FALSE,
                      size = "Huge")
```

# Objectives

This section covers some packages for displaying data in tables in your R Markdown reports. 

# Resources 

Slides: this is an extra lesson, so there aren't currently slides available (yet!)

RStudio.Cloud: this is an extra lesson, so there isn't an RStudio.Cloud project (yet!)

# Packages

We'll be covering a few different packages to use for displaying data in tables: 

+ [`gt`](https://gt.rstudio.com/)   
+ [`reactable`](https://glin.github.io/reactable/)  
+ [`DT`](https://rstudio.github.io/DT/)  

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(gt)
library(reactable)
library(DT)
```


# Data

The data for this section comes from the `fivethirtyeight` and `tidytuesdayR` packages. 

```{r data-package, message=FALSE, warning=FALSE}
library(fivethirtyeight)
library(tidytuesdayR)
```

## 


```{r}
tuesdata <- tidytuesdayR::tt_load('2018-07-03')
```

