---
title: "Displaying Tables With reactable (under construction)"
comment: "*interactive tables in reproducible reports*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: hide
    theme: flatly
    df_print: paged

always_allow_html: true
---

```{r , include=FALSE}
library(tidyverse)
library(skimr)
# create data folder
# folders
folders <- c("code", "data", "img", "docs", "pdfs")
purrr::map(.x = folders, .f = fs::dir_create)
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      tidy = FALSE,
                      size = "Huge")
```

# Objectives

This section covers some packages for displaying data in tables in your R Markdown reports. 

# Resources 

Slides: this is an extra lesson, so there aren't currently slides available (yet!)

RStudio.Cloud: this is an extra lesson, so there isn't an RStudio.Cloud project (yet!)

# Packages

We'll be covering the [`reactable`](https://glin.github.io/reactable/) package for displaying tables in R Markdown.
+   
+ [`DT`](https://rstudio.github.io/DT/)  

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(gt)
library(reactable)
library(DT)
```

# Yahoo finance data

We're going to be returning to stock data for this section. Stock (or quant) data is typically presented on the web or in a dashboard, so it's great for demonstrating the capabilities of R table packages. 

Here is the source for the dataset below: https://finance.yahoo.com/u/yahoo-finance/watchlists/biotech-and-drug-stocks

## Example summary table  {.tabset}

These data have already been formatted in a way that is great for a table, but would pose problems if we tried to visualize them. For example, the `percent_change`-`market_cap` columns have symbols that cause them to be read into R as characters. We would need to wrangle these data quite a bit to get them into a format we could use.

```{r biotech-and-drug-stocks, message=FALSE, warning=FALSE}
BioTechStocksRaw <- read_csv("data/2020-12-20-BioTechStocks.csv")
BioTechStocksRaw
```

```{r BioTechStocksAll, message=FALSE, warning=FALSE}
BioTechStocksRaw %>%
  mutate(
    # change numeric
    change_num = str_remove_all(string = change, pattern = "^\\+"),
    change_num = as.numeric(change_num, na.rm = TRUE),
    # percent change
    perc_change_num = str_remove_all(string = percent_change, pattern = "^\\+"),
    perc_change_num = str_remove_all(string = perc_change_num, pattern = "\\%"),
    perc_change_num = as.numeric(perc_change_num, na.rm = TRUE),
    # market cap
    market_cap_b = str_remove_all(string = market_cap, pattern = "B"), 
    market_cap_b = as.numeric(market_cap_b, na.rm = TRUE)) %>% 
  select(symbol:last_price,
         change_chr = change,
         change_num, 
         perc_change_chr = percent_change,
         perc_change_num,
         volume,
         avg_vol_3_month,
         market_cap_chr = market_cap, 
         market_cap_b,
         market_time) -> BioTechStocksChngeMrktCap
BioTechStocksChngeMrktCap
```

Now we'll reformat the volume and average 3-month volume variables and store the clean variables in `BioTechStocksClean`.

```{r BioTechStocksClean}
BioTechStocksChngeMrktCap %>% 
  mutate(volume_key = 
           case_when(
    str_detect(string = volume, pattern = "k") ~ "volume ($K)", 
    str_detect(string = volume, pattern = "M") ~ "volume ($M)"),
    avg_vol_3mo_key = 
      case_when(
    str_detect(string = avg_vol_3_month, pattern = "k") ~ "volume ($K)", 
    str_detect(string = avg_vol_3_month, pattern = "M") ~ "volume ($M)"),
    # volume
    volume_chr = str_remove_all(string = volume, pattern = "M|k"), 
    volume_chr = as.numeric(volume_chr, na.rm = TRUE),
    # average vol 3-month
    avg_vol_3mo_chr = str_remove_all(string = avg_vol_3_month, pattern = "M|k"), 
    avg_vol_3mo_chr = as.numeric(avg_vol_3mo_chr, na.rm = TRUE),
    # format volume
    volume_num = case_when(
      volume_key == "volume ($K)" ~ round(volume_chr*0.001, digits = 2),
      volume_key == "volume ($M)" ~ round(volume_chr, digits = 2)),
    avg_vol_3mo_num = case_when(
      avg_vol_3mo_key == "volume ($K)" ~ round(avg_vol_3mo_chr*0.001, digits = 2),
      avg_vol_3mo_key == "volume ($M)" ~ round(avg_vol_3mo_chr, digits = 2))) %>% 
  dplyr::select(Symbol = symbol, 
                Company = company_name, 
                `Last Price` = last_price, 
                Change = change_num, 
                `Percent Change` = perc_change_num, 
                `Market Time` = market_time,
                `Volume (Mil)` = volume_num,
                `Avg 3month Volume (Mil)` = avg_vol_3mo_num, 
                `Market Cap (Bil)` = market_cap_b) -> BioTechStocksClean
BioTechStocksClean
```

## Stock data from `quantmod` {.tabset}

We will also import some actual stock data using `quantmod` and `xts` packages.

```{r biotech_symbols, message=FALSE, warning=FALSE}
source("code/quant-packages.R")
biotech_symbols <- c("BMRN", "NVO", "SGEN", "SNY", "BMY")
biotech_prices <- quantmod::getSymbols(biotech_symbols, 
                  # returns an object with the opening price, closing price, 
                  # adjusted price, daily high, daily low and daily volume
                  # for each ticker.
                     src = 'yahoo', 
                     from = "2019-12-31", to = "2020-12-20", 
                     auto.assign = TRUE, 
                     warnings = FALSE) %>% 
          # Ad() and get() is used to isolate the adjusted price
          purrr::map(~Ad(get(.))) %>% 
          # merge() looks for the date index shared by our objects and 
          # uses that index to align the data.
          purrr::reduce(merge) %>% 
          # use colnames<- to rename the columns according to the 
          # symbols object.
          `colnames<-`(biotech_symbols)
```

### Convert adjusted prices to wide 

Here is the wide version of the prices.

```{r PricesWide}
biotech_prices %>%
  xts::to.monthly(indexAt = "lastof", 
             OHLC = FALSE) %>%
  # convert the index to a date
  base::data.frame(date = index(.)) %>%
  # now remove the index because it got converted to row names 
  tibble::remove_rownames() %>% 
  dplyr::select(date, 
                dplyr::everything()) -> PricesWide
PricesWide
```

### Calculate log returns wide

Below we create a wide dataset of log returns. 

```{r ReturnsWide}
PricesWide %>%
  # shift to long 
  tidyr::pivot_longer(names_to = "asset", 
                      values_to = "prices", 
                      cols = -date) %>% 
  dplyr::group_by(asset) %>% 
  # calculate log returns and add those returns to the data frame
  dplyr::mutate(returns = (log(prices) - log(lag(prices)))) %>% 
  dplyr::select(-prices) %>% 
  tidyr::pivot_wider(names_from = asset, 
                     values_from = returns) %>% 
  dplyr::filter(!is.na(BMRN)) -> ReturnsWide
ReturnsWide
```

## Tidy adjusted prices & log returns {.tabset}

Below we filter both wide datasets to `2020-12-31`, pivot them into a long format, and join them into a single dataset with `Symbols`, `Adj Prices`, and `Log Returns`.

### Prices 2020-12

```{r bt_price_12_2020}
bt_prices_12_2020 <- PricesWide %>% 
  filter(date == "2020-12-31") %>% 
  pivot_longer(names_to = "symbols", 
               values_to = "prices", cols = -date)
bt_prices_12_2020
```

### Returns 2020-12

```{r bt_returns_12_2020}
bt_returns_12_2020 <- ReturnsWide %>% 
  filter(date == "2020-12-31") %>% 
  pivot_longer(names_to = "symbols", 
               values_to = "returns", cols = -date)
bt_returns_12_2020
```

### Adj Prices & Log Returns

```{r bt_stocks_12_2020}
bt_stocks_12_2020 <- bt_prices_12_2020 %>% 
  inner_join(x = ., 
           y = bt_returns_12_2020, 
           by = c("date", "symbols")) %>% 
  mutate(returns = round(returns, digits = 4),
         prices = round(prices, digits = 4)) %>% 
  rename(`Adj Prices` = prices,
         `Log Returns` = returns, 
         Symbols = symbols,
         Date = date)
bt_stocks_12_2020
```

# `reactable` package

The `reactable` package provides a comprehensive library of data display features. 

## Dealing with symbols {.tabset}

Adding `+` and `-` to a table (like the )

### adding symbols 

```{r reactable-colDef}
reactable(bt_stocks_12_2020, 
          columns = list( # add + or - for returns
            `Log Returns` = colDef( # define cell function
                      cell = function(value) { 
                      # here we use paste0() to combine the value with +
                      # for all values greater than 0
                       if (value >= 0) paste0("+", value) 
                       else value
                      }, # add color for values above or below 0
                      style = function(value) {
                              color <- if (value > 0) { 
                              # green if above 
                              "#008000"
                              # red if below
                              } else if (value < 0) {
                              "#e00000"
                              } # adjust the font weight and color
                  list(fontWeight = 600, color = color)
                  }
                )
            )
          )
```






