---
title: "Displaying Tables With gt (under construction)"
comment: "*interactive tables in reproducible reports*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: hide
    theme: flatly
    df_print: paged

always_allow_html: true
---

```{r , include=FALSE}
library(tidyverse)
library(skimr)
# create data folder
# folders
folders <- c("code", "data", "img", "docs", "pdfs")
purrr::map(.x = folders, .f = fs::dir_create)
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      tidy = FALSE,
                      size = "Huge")
```

# Objectives

This section covers some packages for displaying data in tables in your R Markdown reports.

# Resources

Slides: this is an extra lesson, so there aren't currently slides available (yet!)

RStudio.Cloud: this is an extra lesson, so there isn't an RStudio.Cloud project (yet!)

# Packages

We'll be covering the [`gt`](https://gt.rstudio.com/) package for displaying tables in R Markdown.

# Stock data from `quantmod` 

We will also import some actual stock data using `quantmod` and `xts` packages.

```{r biotech_symbols, message=FALSE, warning=FALSE}
source("code/quant-packages.R")
biotech_symbols <- c("BMRN", "NVO", "SGEN", "SNY", "BMY")
biotech_prices <- quantmod::getSymbols(biotech_symbols, 
                  # returns an object with the opening price, closing price, 
                  # adjusted price, daily high, daily low and daily volume
                  # for each ticker.
                     src = 'yahoo', 
                     from = "2019-12-31", to = "2020-12-31", 
                     auto.assign = TRUE, 
                     warnings = FALSE) %>% 
          # Ad() and get() is used to isolate the adjusted price
          purrr::map(~Ad(get(.))) %>% 
          # merge() looks for the date index shared by our objects and 
          # uses that index to align the data.
          purrr::reduce(merge) %>% 
          # use colnames<- to rename the columns according to the 
          # symbols object.
          `colnames<-`(biotech_symbols)
```

## Convert adjusted prices to wide {.tabset}

Here is the wide version of the prices.

### `PricesWide`

```{r PricesWide}
biotech_prices %>%
  xts::to.monthly(indexAt = "lastof", 
             OHLC = FALSE) %>%
  # convert the index to a date
  base::data.frame(date = index(.)) %>%
  # now remove the index because it got converted to row names 
  tibble::remove_rownames() %>% 
  dplyr::select(date, 
                dplyr::everything()) -> PricesWide
PricesWide
```

### `ReturnsWide`

Below we create a wide dataset of log returns. 

```{r ReturnsWide}
PricesWide %>%
  # shift to long 
  tidyr::pivot_longer(names_to = "asset", 
                      values_to = "prices", 
                      cols = -date) %>% 
  # 
  dplyr::group_by(asset) %>% 
  # calculate log returns and add those returns to the data frame
  dplyr::mutate(returns = (log(prices) - log(lag(prices)))) %>% 
  dplyr::select(-prices) %>% 
  tidyr::pivot_wider(names_from = asset, 
                     values_from = returns) %>% 
  dplyr::filter(!is.na(BMRN)) -> ReturnsWide
ReturnsWide
```

## Tidy adjusted prices & log returns {.tabset}

Below we filter both wide datasets to `2020-12-31`, pivot them into a long format, and join them into a single dataset with `Symbols`, `Adj Prices`, and `Log Returns`.

### Prices 2020-12

```{r bt_price_12_2020}
bt_prices_12_2020 <- PricesWide %>% 
  filter(date == "2020-12-31") %>% 
  pivot_longer(cols = -date, 
               names_to = "symbols", 
               values_to = "prices")
bt_prices_12_2020
```

### Returns 2020-12

```{r bt_returns_12_2020}
bt_returns_12_2020 <- ReturnsWide %>% 
  filter(date == "2020-12-31") %>% 
  pivot_longer(names_to = "symbols", 
               values_to = "returns", cols = -date)
bt_returns_12_2020
```

### Adj Prices & Log Returns

```{r bt_stocks_12_2020}
bt_stocks_12_2020 <- bt_prices_12_2020 %>% 
  inner_join(x = ., 
           y = bt_returns_12_2020, 
           by = c("date", "symbols")) %>% 
  mutate(returns = round(returns, digits = 4),
         prices = round(prices, digits = 4)) %>% 
  rename(`Adj Prices` = prices,
         `Log Returns` = returns, 
         Symbols = symbols,
         Date = date)
bt_stocks_12_2020
```

# `gt` package

## Dealing with merged columns {.tabset}

### merging columns 

Merge columns using...

### column headers 

Add column headers using...
