---
title: "Introduction to Dashboards (under construction)"
comment: "*building data displays with `flexdashboard`*"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
    number_sections: yes
    code_folding: show
    theme: flatly
    df_print: paged

always_allow_html: true
---


```{r , include=FALSE}
library(tidyverse)
library(skimr)
library(lubridate)
library(here)
library(janitor)
library(flexdashboard)
library(ggrepel)
library(covdata)
library(showtext)
library(hrbrthemes)
# create data folder
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.path = "img/",
                      tidy = FALSE)
# set width
options(width = 60, max.print = 60)
```

# Objectives

An introduction to dashboard development in R with `flexdashboard`s. 

# Materials

- Slides: https://mjfrigaard.github.io/intro-to-dashboards/Index.html 

- RStudio Project: https://rstudio.cloud/project/2000287

# Previous lessons

All of the exercises and lessons are available [here](https://mjfrigaard.github.io/r-meetup-tutorials/), but you can also read more about [`ggplot2`](https://ggplot2.tidyverse.org/index.html) on the tidyverse website, and in the [Data Visualisation](https://r4ds.had.co.nz/data-visualisation.html) chapter of R for Data Science.

# Load the `tidyverse` and `flexdashboard`

The main packages we're going to use are `dplyr`, `tidyr`, `ggplot2`, and `flexdashboard`. 


```{r packages, eval=FALSE}
library(tidyverse)
library(skimr)
library(janitor)
library(flexdashboard)
```

# What should go into a dashboard?
